---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
editor_options: 
  markdown: 
    wrap: 80
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(GGally)
library(corrplot)
library(olsrr)
```

--------------------------------------------------------------------------------

## Part 1: Data

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit
your work.

```{r load-data}
load("movies.Rdata")
```

The data set is comprised of 651 randomly sampled movies produced and released
before 2016.

Some of these variables are only there for informational purposes and do not
make any sense to include in a statistical analysis. It is up to you to decide
which variables are meaningful and which should be omitted. For example
information in the the `actor1` through `actor5` variables was used to determine
whether the movie casts an actor or actress who won a best actor or actress
Oscar.

You might also choose to omit certain observations or restructure some of the
variables to make them suitable for answering your research questions.

When you are fitting a model you should also be careful about collinearity, as
some of these variables may be dependent on each other.

### Codebook

1.  `title`: Title of movie

2.  `title_type`: Type of movie (Documentary, Feature Film, TV Movie)

3.  `genre`: Genre of movie (Action & Adventure, Comedy, Documentary, Drama,
    Horror, Mystery & Suspense, Other)

4.  `runtime`: Runtime of movie (in minutes)

5.  `mpaa_rating`: MPAA rating of the movie (G, PG, PG-13, R, Unrated)

6.  `studio`: Studio that produced the movie

7.  `thtr_rel_year`: Year the movie is released in theaters

8.  `thtr_rel_month`: Month the movie is released in theaters

9.  `thtr_rel_day`: Day of the month the movie is released in theaters

10. `dvd_rel_year`: Year the movie is released on DVD

11. `dvd_rel_month`: Month the movie is released on DVD

12. `dvd_rel_day`: Day of the month the movie is released on DVD

13. `imdb_rating`: Rating on IMDB

14. `imdb_num_votes`: Number of votes on IMDB

15. `critics_rating`: Categorical variable for critics rating on Rotten Tomatoes
    (Certified Fresh, Fresh, Rotten)

16. `critics_score`: Critics score on Rotten Tomatoes

17. `audience_rating`: Categorical variable for audience rating on Rotten
    Tomatoes (Spilled, Upright)

18. `audience_score`: Audience score on Rotten Tomatoes

19. `best_pic_nom`: Whether or not the movie was nominated for a best picture
    Oscar (no, yes)

20. `best_pic_win`: Whether or not the movie won a best picture Oscar (no, yes)

21. `best_actor_win`: Whether or not one of the main actors in the movie ever
    won an Oscar (no, yes) – note that this is not necessarily whether the actor
    won an Oscar for their role in the given movie

22. `best_actress win`: Whether or not one of the main actresses in the movie
    ever won an Oscar (no, yes) – not that this is not necessarily whether the
    actresses won an Oscar for their role in the given movie

23. `best_dir_win`: Whether or not the director of the movie ever won an Oscar
    (no, yes) – not that this is not necessarily whether the director won an
    Oscar for the given movie

24. `top200_box`: Whether or not the movie is in the Top 200 Box Office list on
    BoxOfficeMojo (no, yes)

25. `director`: Director of the movie

26. `actor1`: First main actor/actress in the abridged cast of the movie

27. `actor2`: Second main actor/actress in the abridged cast of the movie

28. `actor3`: Third main actor/actress in the abridged cast of the movie

29. `actor4`: Fourth main actor/actress in the abridged cast of the movie

30. `actor5`: Fifth main actor/actress in the abridged cast of the movie

31. `imdb_url`: Link to IMDB page for the movie

32. `rt_url`: Link to Rotten Tomatoes page for the movie

--------------------------------------------------------------------------------

## Part 2: Research question

-   What attributes make a movie popular?
-   anything else new and interesting about movie?

--------------------------------------------------------------------------------

## Part 3: Exploratory data analysis

From research questions above, we need to find the best model that predict
popularity of a movies. Given the cookbook of the dataset, we can see the
variable **imdb_num_votes** is the strongest variable that can indicates the
popularity of a movies. Other variables contained score and rating just show the
average score/rating of a movie out of a number of voters. While
**imdb_num_votes** tells us how many people know and rate the film.

Thus we make **imdb_num_votes** to be the response variable of our model. the
more number of votes a movie receives on imdb platforms, the more popular that
movie is.

### Clean the dataset

-   Some variables are just information that won't predict anything, so we
    cannot use in them our analysis:

    -   movie title

    -   director

    -   actor1→5

    -   imdb_url, rt_url

```{r}
# read the movies table
glimpse(movies)
# remove non-predicted variable
df <- movies[,-c(1,6, 32, 31, 30, 29, 28, 27, 26, 25, 24)]
```

### Examine response variable

```{r}
# analyze the response var
summary(df$imdb_num_votes)
boxplot(df$imdb_num_votes)
hist(df$imdb_num_votes,breaks = 200)
```

Finding

-   The distribution of popularity of movies is extremely right skewed

### Examine "released time"

We examine "released time" variables to find out their relationship with the
model

```{r}
plot(x = movies$thtr_rel_year, y = movies$imdb_num_votes)
plot(x = movies$thtr_rel_month, y = movies$imdb_num_votes)
plot(x = movies$thtr_rel_day, y = movies$imdb_num_votes)
plot(x = movies$dvd_rel_year, y = movies$imdb_num_votes)
plot(x = movies$dvd_rel_month, y = movies$imdb_num_votes)
plot(x = movies$dvd_rel_day, y = movies$imdb_num_votes)
```

Finding:

-   The years when a movie is released, both in theaters and on DVD, often
    indicate its popularity. Movies released more recently tend to receive more
    ratings.

```{=html}
<!-- -->
```
-   The specific month and day when a movie is released in theaters or on DVD
    don't provide useful information.

-\> We'll make new variables to store how long ago a movie was released in
theaters and on DVD. This will help us analyze their relationship in the model
later

```{r}
# add "age" variables to the data
df <- df %>% mutate(thtr_age = 2016 - thtr_rel_year, 
              dvd_age = 2016 - dvd_rel_year)
# plot
plot(y = df$imdb_num_votes, x = df$dvd_age)
plot(y = df$imdb_num_votes, x = df$thtr_age)

# remove "released time" variables
df <- df[, -c(5,6,7,8,9,10)]
```

### Examine categorical variables

##### title_type, genre, mpaa_rating

```{r}
df_sub1 <- subset(df, select = c(imdb_num_votes, title_type, genre, mpaa_rating))
ggpairs(df_sub1) 
```

##### critics_rating, audience_rating

```{r}
df_sub2 <- subset(df, select = c(imdb_num_votes, critics_rating, audience_rating))
ggpairs(df_sub2) 
```

##### best_pic_nom, best_pic_win, best_actor_win, best_actress_win, best_dir_win

```{r}
df_sub3 <- subset(df, select = c(imdb_num_votes, best_pic_nom, best_pic_win, best_actor_win, best_actress_win, best_dir_win))
ggpairs(df_sub3) 
```

### Examine others numerical variables

```{r}
df_num <- df[, c(3,5,6,8,10)]
m1 = cor(df_num)
corrplot(m1, method = 'number')
ggpairs(df_num) 
```

Finding:

-   run_time didn't show any significant relationship with any other variables →
    we can drop this variable from the dataset.

-   imdb_rating, audience_score and critics_scores are highly correlated → we
    should remove some variables to avoid collinearity. My suggestion is to
    remove the imdb_rating variable, because the others have lower correlation
    score. Additionally, audience and critics scores offer distinct
    perspectives, which adds valuable insight.

```{r}
df <- df[, -c(3,5)]
```

--------------------------------------------------------------------------------

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk"
button above. Make sure that your code is visible in the project you submit.
Delete this note when before you submit your work.

```{r}
m_full <- lm(imdb_num_votes ~ title_type + genre + mpaa_rating + critics_rating + critics_score + audience_rating + audience_score + best_pic_nom + best_pic_win + best_actor_win + best_actress_win + best_dir_win + thtr_age + dvd_age, df)
summary(m_full)
```

### Model Selection

We use backward elimination based on R-squared

1.  We remove some variables with very high insignificant p-value

```{r}
m_full_1 <- lm(imdb_num_votes ~ genre + critics_rating + critics_score + audience_rating + audience_score + best_pic_nom + best_pic_win + best_actor_win + best_actress_win + best_dir_win + thtr_age + dvd_age, df)
```

2.  Then use a function from `olsrr` package to test all possible subset of
    explanatory variables to find the best model

```{r}
ols_step_best_subset(m_full_1, metric = "rsquare")
```

2.  Final model

```{r}
m_final <- lm(imdb_num_votes ~ critics_score + audience_rating + audience_score + best_pic_nom + best_pic_win + best_dir_win + thtr_age, df)
summary(m_final)
```

--------------------------------------------------------------------------------

## Part 5: Prediction

Use data on the dataset to predict

```{r}
df_test <- df[1:10,]
predict(m_final, df_test)
```

--------------------------------------------------------------------------------

## Part 6: Conclusion
